---
title: "Observed acquisition integrity: hardware, SDK and analysis clocks"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Observed acquisition integrity}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The 2026 methodological briefings distinguish hardware-native capability,
what the acquisition SDK delivered, the observed timestamps and the grid used
by downstream analysis. The distinction matters when low-rate recordings
are interpolated to a much denser computational grid: interpolation does not
recover lost physiological information.

```{r example, eval=requireNamespace('ggplot2', quietly=TRUE)}
library(eyeprocess)
audit <- audit_eye_acquisition_integrity(
  time_s = c(0, .01, .02, .02, .06, .061, .07),
  valid = c(TRUE, TRUE, FALSE, TRUE, TRUE, TRUE, TRUE),
  device_native_rate_hz = 100,
  sdk_delivered_rate_hz = 90,
  analysis_grid_rate_hz = 500
)
audit$summary
plot_eye_acquisition_integrity(audit)
```

The plot retains the zero timestamp interval and long gap. The declared
device/SDK rates are supplied metadata; the helper does **not** verify those
against firmware, infer absent packets, repair clock drift or establish
millisecond-accurate cross-device alignment. The candidate catch-up flag is a
descriptive long-then-short interval heuristic, not a diagnosis of acquisition
software bugs.

This addition is under development; no release is implied.


## DABEST-style effect visualization after eyeprocess preprocessing

Once gaze/pupil observations are reduced to **independent participant-level**
effect summaries, use the official R
[dabestr package](https://acclab.github.io/dabestr/) for Gardner–Altman
or Cumming plots. These deliberately display both the raw distribution
and BCa bootstrap uncertainty. The underlying
[DABEST Python examples](https://acclab.github.io/DABEST-python/)
are the visualization reference; no cross-language code is copied.

```{r dabest-bridge, eval=FALSE}
# aggregated <- one validated row per participant and condition,
# with ParticipantID, Condition and PupilChange columns.
library(dabestr)
d <- dabestr::load(
  aggregated, x = Condition, y = PupilChange,
  idx = c("low", "high"),
  paired = "baseline", id_col = ParticipantID
)
dabestr::dabest_plot(dabestr::mean_diff(d), float_contrast = FALSE)
```

Sample-level eye-tracking data, event windows and serially correlated
trials cannot be resampled as independent observations. For those designs,
retain the package's grouped/multilevel inference rather than substituting
unadjusted raw mean differences.
