---
title: "Adaptive testing design and governance"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Adaptive testing design and governance}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(eyeprocess)
```

Native adaptive-design utilities deliberately remain simple and auditable: item-bank declaration, maximum-information selection, exposure summaries, content-balance audits, stopping rules, and step-by-step adaptive traces.

```{r}
items <- data.frame(item_id=paste0("I",1:6),a=c(.8,1,1.5,1.1,.9,1.2),b=seq(-1.5,1.5,length.out=6),c=0,d=1)
bank <- eyeprocess_irt_item_bank(items, content = c("A","A","B","B","C","C"))
eyeprocess_irt_item_selection(bank, theta = 0)
eyeprocess_irt_stopping_rule(10, se = .25)
```

For mature CAT simulation and constrained/shadow-testing designs, eyeprocess delegates to `catR` and `mirtCAT`. Availability is explicit and no simpler CAT engine is silently substituted.

Primary package sources: <https://cran.r-project.org/package=catR> and <https://cran.r-project.org/package=mirtCAT>.
