---
title: "AOI Perturbation and Uncertainty Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AOI Perturbation and Uncertainty Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4.5)
library(eyeprocess)
```

## AOI Geometry Is an Analytical Assumption

AOI geometry is an analytical assumption. Small changes to a boundary can
change whether a sample or fixation is counted as headline, image, claim,
disclosure, CTA, outside, or ambiguous. The purpose of this workflow is not to
find a preferred boundary. It asks whether a conclusion is stable under a
declared set of defensible alternatives.

Use this analysis when observations cluster near boundaries, neighboring AOIs
are close, or AOI size is comparable with expected spatial error. Do not use it
as a substitute for calibration, coordinate-quality auditing, or preregistered
stimulus design.

## Synthetic advertising layout

The example is fully synthetic and intentionally small.

```{r data}
aois <- data.frame(
  aoi_id = c("headline", "image", "claim", "disclosure", "cta"),
  xmin = c(100, 300, 300, 300, 680),
  xmax = c(400, 700, 700, 700, 920),
  ymin = c(80, 190, 360, 500, 600),
  ymax = c(180, 340, 480, 590, 710)
)

set.seed(12)
centers <- data.frame(
  x = c(250, 500, 500, 500, 800),
  y = c(130, 270, 430, 560, 650),
  aoi = c("headline", "image", "claim", "disclosure", "cta")
)

rows <- list()
k <- 0L
for (participant in 1:8) {
  condition <- participant %% 2
  for (trial in 1:3) {
    for (j in seq_len(nrow(centers))) {
      n <- 5L + if (centers$aoi[j] == "disclosure" && condition == 1) 3L else 0L
      for (fix in seq_len(n)) {
        k <- k + 1L
        rows[[k]] <- data.frame(
          obs = k,
          participant = paste0("p", participant),
          trial = trial,
          condition = condition,
          x = rnorm(1, centers$x[j], 35),
          y = rnorm(1, centers$y[j], 22),
          duration = runif(1, .06, .18),
          time = j - 1 + (fix - 1) / 20
        )
      }
    }
  }
}
fixations <- do.call(rbind, rows)
```

## AOI Perturbation Sensitivity Analysis\n\n### Pixels versus Degrees of Visual Angle

Degree-based perturbations require explicit display geometry and viewing
distance. The workflow never silently converts units.

```{r grid}
grid <- create_aoi_perturbation_grid(
  dilations = c(.25, .50, 1.00),
  erosions = .25,
  translations_x = .50,
  translations_y = .50,
  unit = "deg",
  include_baseline = TRUE,
  screen_width_px = 1024,
  screen_height_px = 768,
  viewing_distance = 60,
  physical_screen_size = c(53.1, 29.9),
  boundary_policy = "allow"
)

grid$table
```

The seven branches are the nominal geometry, three dilations, one erosion,
one horizontal translation, and one vertical translation.

## Propagating AOI Uncertainty Into Statistical Models

The perturbation framework does not choose an estimator. Here the example
uses the same base-R linear model for disclosure dwell in every branch.

```{r callback}
model_callback <- function(features, assigned, spec) {
  target <- features[features$aoi == "disclosure", , drop = FALSE]
  condition_map <- unique(assigned[, c("participant", "condition")])
  target <- merge(target, condition_map, by = "participant", all.x = TRUE)

  fit <- stats::lm(dwell ~ condition, data = target)
  co <- summary(fit)$coefficients
  if (!"condition" %in% rownames(co)) stop("condition coefficient unavailable")

  estimate <- unname(co["condition", "Estimate"])
  se <- unname(co["condition", "Std. Error"])

  data.frame(
    term = "condition",
    estimate = estimate,
    SE = se,
    CI_low = estimate - 1.96 * se,
    CI_high = estimate + 1.96 * se,
    p_value = unname(co["condition", "Pr(>|t|)"]),
    model_converged = TRUE,
    N = stats::nobs(fit)
  )
}
```

## Run geometry to inference

```{r run}
result <- run_aoi_sensitivity_analysis(
  fixations,
  aois,
  grid,
  x_col = "x",
  y_col = "y",
  observation_id_col = "obs",
  participant_col = "participant",
  trial_col = "trial",
  duration_col = "duration",
  time_col = "time",
  overlap_policy = "ambiguous",
  model_callback = model_callback,
  preprocessing_specification = list(duration_unit = "seconds"),
  event_detector = "synthetic_fixations",
  quality_rules = list(missing = "preserve", overlap = "ambiguous"),
  model_specification = list(
    family = "lm",
    outcome = "disclosure_dwell",
    predictor = "condition"
  )
)

summary <- summarise_aoi_sensitivity(result)
summary$assignment_stability
summary$inference_stability
```

## Interpreting AOI Assignment Stability\n\nThe assignment table distinguishes unchanged observations, newly assigned\nobservations, lost assignments, AOI-to-AOI reassignment, missing comparisons,\nand explicit overlap ambiguity. Missing coordinates are not converted to\noutside-AOI observations.\n
## Visual diagnostics

```{r geometry-plot}
plot_aoi_perturbations(
  result,
  perturbation_id = "dilate_0.5_deg",
  data = fixations[seq(1, nrow(fixations), by = 12), ],
  x_col = "x",
  y_col = "y"
)
```

```{r assignment-plot}
plot_aoi_assignment_stability(result)
```

```{r coefficient-plot}
plot_aoi_coefficient_stability(result, "condition")
```

The coefficient plot should be interpreted together with the assignment plot.
A result can have many reassigned fixations without reversing a coefficient, or
can show high aggregate assignment stability while a model estimate remains
sensitive.

## Demonstrate a failure branch

Excessive erosion is retained as a failed branch rather than disappearing.

```{r failure}
failure_grid <- create_aoi_perturbation_grid(erosions = 500)
failure_result <- apply_aoi_perturbation_grid(aois, failure_grid)
failure_result$audit
```

Model callbacks follow the same rule: an error is written to the failure table,
and a row with `model_converged = FALSE` remains visible rather than being
treated as a valid estimate.

## Reporting AOI Robustness

```{r report}
cat(report_aoi_sensitivity(result))
```

A manuscript should report the nominal AOI source, coordinate system,
perturbation operations and values, pixel-versus-degree units, screen geometry
and viewing distance when degrees are used, overlap policy, screen-boundary
policy, random seed for jitter, assignment level, recomputed features, model
specification, and handling of failed or non-converged branches.

Robustness frequencies are descriptive summaries conditional on the declared
perturbation set. They are **not probabilities that the scientific conclusion
is true**.

## Limitations

- Geometry perturbation does not replace calibration-error analysis.
- It does not replace event-detector sensitivity analysis.
- Exact concave-polygon buffering requires a dedicated geometry engine; the
  dependency-free core rejects concave dilation/erosion rather than silently
  approximating it.
- Degree conversion inherits uncertainty in screen dimensions and viewing
  distance.
- The perturbation envelope must be scientifically defensible; wider grids do
  not automatically provide stronger evidence.
