---
title: "AOI Perturbation and Uncertainty Analysis"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{AOI Perturbation and Uncertainty Analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", fig.width = 7, fig.height = 4.5)
library(eyeprocess)
```

## AOI Geometry Is an Analytical Assumption

AOI geometry is an analytical assumption. Small changes to a boundary can
change whether a sample or fixation is counted as headline, image, claim,
disclosure, CTA, outside, or ambiguous. The purpose of this workflow is not to
find a preferred boundary. It asks whether a conclusion is stable under a
declared set of defensible alternatives.

Use this analysis when observations cluster near boundaries, neighboring AOIs
are close, or AOI size is comparable with expected spatial error. Do not use it
as a substitute for calibration, coordinate-quality auditing, or preregistered
stimulus design.

## Synthetic advertising layout

The example is fully synthetic and intentionally small.

```{r data}
aois <- data.frame(
  aoi_id = c("headline", "image", "claim", "disclosure", "cta"),
  xmin = c(100, 300, 300, 300, 680),
  xmax = c(400, 700, 700, 700, 920),
  ymin = c(80, 190, 360, 500, 600),
  ymax = c(180, 340, 480, 590, 710)
)

set.seed(12)
centers <- data.frame(
  x = c(250, 500, 500, 500, 800),
  y = c(130, 270, 430, 560, 650),
  aoi = c("headline", "image", "claim", "disclosure", "cta")
)

rows <- list()
k <- 0L
for (participant in 1:8) {
  condition <- participant %% 2
  for (trial in 1:3) {
    for (j in seq_len(nrow(centers))) {
      n <- 5L + if (centers$aoi[j] == "disclosure" && condition == 1) 3L else 0L
      for (fix in seq_len(n)) {
        k <- k + 1L
        rows[[k]] <- data.frame(
          obs = k,
          participant = paste0("p", participant),
          trial = trial,
          condition = condition,
          x = rnorm(1, centers$x[j], 35),
          y = rnorm(1, centers$y[j], 22),
          duration = runif(1, .06, .18),
          time = j - 1 + (fix - 1) / 20
        )
      }
    }
  }
}
fixations <- do.call(rbind, rows)
```

## AOI Perturbation Sensitivity Analysis

### Pixels versus Degrees of Visual Angle

Degree-based perturbations require explicit display geometry and viewing
distance. The workflow never silently converts units.

```{r grid}
grid <- create_aoi_perturbation_grid(
  dilations = c(.25, .50, 1.00),
  erosions = .25,
  translations_x = .50,
  translations_y = .50,
  unit = "deg",
  include_baseline = TRUE,
  screen_width_px = 1024,
  screen_height_px = 768,
  viewing_distance = 60,
  physical_screen_size = c(53.1, 29.9),
  boundary_policy = "allow"
)

grid$table
```

The seven branches are the nominal geometry, three dilations, one erosion,
one horizontal translation, and one vertical translation.

## Propagating AOI Uncertainty Into Statistical Models

The perturbation framework does not choose an estimator. The model callback must explicitly report its estimator output, convergence flag, and model `N`; converged rows require finite estimates, standard errors, confidence limits, and a positive integer-valued `N`. Here the example uses the same base-R linear model for disclosure dwell in every branch.

```{r callback}
model_callback <- function(features, assigned, spec) {
  target <- features[features$aoi == "disclosure", , drop = FALSE]
  condition_map <- unique(assigned[, c("participant", "condition")])
  target <- merge(target, condition_map, by = "participant", all.x = TRUE)

  fit <- stats::lm(dwell ~ condition, data = target)
  co <- summary(fit)$coefficients
  if (!"condition" %in% rownames(co)) stop("condition coefficient unavailable")

  estimate <- unname(co["condition", "Estimate"])
  se <- unname(co["condition", "Std. Error"])

  data.frame(
    term = "condition",
    estimate = estimate,
    SE = se,
    CI_low = estimate - 1.96 * se,
    CI_high = estimate + 1.96 * se,
    p_value = unname(co["condition", "Pr(>|t|)"]),
    model_converged = TRUE,
    N = stats::nobs(fit)
  )
}
```

## Run geometry to inference

```{r run}
result <- run_aoi_sensitivity_analysis(
  fixations,
  aois,
  grid,
  x_col = "x",
  y_col = "y",
  observation_id_col = "obs",
  participant_col = "participant",
  trial_col = "trial",
  duration_col = "duration",
  time_col = "time",
  observation_level = "fixation",
  overlap_policy = "ambiguous",
  model_callback = model_callback,
  preprocessing_specification = list(duration_unit = "seconds"),
  event_detector = "synthetic_fixations",
  quality_rules = list(missing = "preserve", overlap = "ambiguous"),
  model_specification = list(
    family = "lm",
    outcome = "disclosure_dwell",
    predictor = "condition"
  )
)

summary <- summarise_aoi_sensitivity(result)
summary$assignment_stability
summary$inference_stability
```

## Interpreting AOI Assignment Stability

The assignment table distinguishes unchanged observations, newly assigned
observations, lost assignments, AOI-to-AOI reassignment, missing comparisons,
and explicit overlap ambiguity. Missing coordinates are not converted to
outside-AOI observations.



### Fixation and sample input

The same geometry and assignment engine supports fixation-centroid and sample-level input. Declare `observation_level = "fixation"` or `"sample"`. Output always contains universal observation counts/timing; fixation-specific fields are populated only for fixation input and sample counts only for sample input. This prevents sample rows from being reported as fixation counts.

### Zero is not missing

Feature recomputation preserves the complete declared AOI universe within each observed participant/trial group. A zero `observation_count` means valid observations were available but none were assigned to that AOI; `fixation_count` or `sample_count` is populated only at the matching observation level. If a group has no valid AOI assignment opportunity, counts and inspection remain missing rather than being coerced to zero. Missing assigned durations make dwell missing, and first-observation timing is never inferred from row order.

## Visual diagnostics

```{r geometry-plot}
plot_aoi_perturbations(
  result,
  perturbation_id = "dilate_0.5_deg",
  data = fixations[seq(1, nrow(fixations), by = 12), ],
  x_col = "x",
  y_col = "y"
)
```

```{r assignment-plot}
plot_aoi_assignment_stability(result)
```

```{r coefficient-plot}
plot_aoi_coefficient_stability(result, "condition")
```

The coefficient plot should be interpreted together with the assignment plot.
A result can have many reassigned fixations without reversing a coefficient, or
can show high aggregate assignment stability while a model estimate remains
sensitive.

## Demonstrate a failure branch

Excessive erosion is retained as a failed branch rather than disappearing.

```{r failure}
failure_grid <- create_aoi_perturbation_grid(erosions = 500)
failure_result <- apply_aoi_perturbation_grid(aois, failure_grid)
failure_result$audit
```

Model callbacks follow the same rule: an error is written to the failure table,
and a row with `model_converged = FALSE` remains visible rather than being
treated as a valid estimate.

## AOI Robustness Decision Clinic

Interpret assignment and model stability together rather than treating one metric as a verdict.

| Pattern | Interpretation | Recommended response |
|---|---|---|
| High assignment stability + stable coefficient | Stable within the declared envelope | Report the envelope and stability summaries; do not call this proof that the nominal AOI is correct |
| Low assignment stability + stable coefficient | Mapping changes but the model-level result is insensitive | Report reassignment matrices and model stability together |
| High assignment stability + unstable coefficient | A small subset of groups or observations may be influential | Inspect participant/trial/AOI summaries, intervals, convergence, and model `N` |
| Low assignment stability + unstable coefficient | The substantive result depends materially on AOI geometry | Describe the nominal result as geometry-sensitive |
| Failed branches | Part of the planned sensitivity design is non-evaluable | Retain failures and explain them; never shrink the denominator silently |

Choose perturbation magnitudes from acquisition precision, display geometry, stimulus layout, or a preregistered rationale. Do not search for the margin that restores statistical significance.

### Reporting example

> We reran AOI assignment, feature extraction, and the prespecified model across the nominal geometry and defensible perturbations. We report reassignment patterns, coefficient ranges and intervals, convergence, and model-N variation. Failed branches remain in the audit trail. Stability frequencies describe robustness to the declared perturbation set and are not probabilities that the scientific conclusion is true.

## Planning the sensitivity analysis before results

For a fillable preregistration-style template and API mapping, see the companion **AOI Sensitivity Analysis Plan and Reporting Template** article.


Record the AOI sensitivity plan before examining branch-specific model results. At minimum, state:

1. the nominal AOI source and geometry representation;
2. whether rows are samples or fixations;
3. the perturbation operations and exact values;
4. pixel or degree units and the display/viewing geometry when degrees are used;
5. overlap and screen-boundary policies;
6. the fixed statistical model specification;
7. how failed geometry, callback errors, and non-convergence will be retained and reported.

Do not widen, shrink, or selectively edit the perturbation grid because a particular branch changes statistical significance.

## Troubleshooting AOI perturbation branches

Diagnose the earliest failing layer first: **geometry -> assignment -> feature recomputation -> model callback**.

| Symptom | Inspect | Interpretation |
|---|---|---|
| Perturbation status is `failed` | `result$grid_result$audit` | Geometry could not be evaluated; keep the branch in the planned denominator |
| Many `__ambiguous__` assignments | perturbed AOIs and reassignment matrix | AOIs overlap under that perturbation; do not silently choose one |
| Many `__outside__` assignments | units, coordinate transform, AOI location | Layout/assignment mismatch, not automatically zero exposure |
| Count is zero | valid-observation denominator | Valid observations existed but none entered that AOI |
| Count is missing | missing-coordinate denominator | No valid assignment opportunity existed; do not coerce to zero |
| Callback failure | `result$failures` | Model branch is non-evaluable, not a valid null result |
| `model_converged = FALSE` | model table and estimator diagnostics | Retain the row but exclude it from converged-effect summaries |
| Model `N` changes | inference stability table | Case loss varies across branches and must be reported |

A compact audit pattern is:

```r
result$grid_result$audit[, c("perturbation_id", "status", "message")]
result$failures
result$models[, c("perturbation_id", "term", "model_converged", "N")]
assess_aoi_inference_stability(result, term = "condition")
```

The package does not reinterpret a geometry failure as a model result, a callback exception as non-significance, or missing assignment opportunity as a structural zero.

### Reporting problematic branches

> Of the planned AOI perturbations, all non-evaluable geometry and model branches were retained in the audit trail. Overlap ambiguity remained explicit, missing assignment opportunities were not converted to zero, and non-converged model rows were excluded from converged-effect summaries. Model-N variation was reported alongside coefficient and assignment stability.

This is a reporting template, not an empirical result.

## Reporting AOI Robustness

For manuscript supplements, reviewer responses, or replication handoff, see the companion **AOI Robustness Reporting Bundle** article for a recommended file set, export pattern, privacy guidance, and API map.


```{r report}
cat(report_aoi_sensitivity(result))
```

A manuscript should report the nominal AOI source, coordinate system, observation level, perturbation operations and values, pixel-versus-degree units, screen geometry and viewing distance when degrees are used, overlap policy, screen-boundary policy, random seed for jitter, assignment denominators, recomputed features, model specification, coefficient/interval ranges, convergence, model-`N` variation, and handling of failed or non-converged branches.

Robustness frequencies are descriptive summaries conditional on the declared
perturbation set. They are **not probabilities that the scientific conclusion
is true**.

## R/Python Scientific Parity

The R and Python implementations share the same perturbation IDs, outside and
ambiguous labels, missing-coordinate semantics, unit requirements, failure
records, model-convergence handling, and interpretation caveat. Both ship the
same frozen two-AOI parity fixture.

Parity is semantic rather than byte-for-byte numerical identity. Base-R and
NumPy random-number generators are allowed to produce different jitter
coordinates from the same integer seed, provided each implementation is
deterministic for its own seed and records that seed. Plot objects are also
language-native: base R graphics in `eyeprocess` and Matplotlib axes in
`eyeprocesspy`.

User-supplied model callbacks may use different statistical backends, so exact
coefficient equality is not required unless the same estimator and numerical
backend are deliberately used.

## Limitations

- Geometry perturbation does not replace calibration-error analysis.
- It does not replace event-detector sensitivity analysis.
- Exact concave-polygon buffering requires a dedicated geometry engine; the
  dependency-free core rejects concave dilation/erosion rather than silently
  approximating it.
- Degree conversion inherits uncertainty in screen dimensions and viewing
  distance.
- The perturbation envelope must be scientifically defensible; wider grids do
  not automatically provide stronger evidence.


## API map

Use `create_aoi_perturbation_grid()` to declare branches, `run_aoi_sensitivity_analysis()` for the complete workflow, `estimate_aoi_assignment_stability()` for assignment robustness, `assess_aoi_inference_stability()` for coefficient/convergence/model-`N` summaries, and `report_aoi_sensitivity()` for a compact reporting draft.
