---
title: "API lifecycle and canonical interfaces"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{API lifecycle and canonical interfaces}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

A large scientific package needs a navigable public surface. The 0.9 lifecycle layer inventories exports, maps them to conceptual families, and permits explicit statuses such as core, workflow, advanced, experimental, gated, compatibility, and deprecated. Unreviewed functions remain `unreviewed`; eyeprocess does not infer maturity from a name alone.

```{r, eval=FALSE}
inv <- eye_api_inventory()
reg <- eye_api_lifecycle()
reg <- register_eye_api_status(reg, "run_eye_pipeline", "workflow", canonical="run_eye_pipeline")
audit <- audit_eye_api(inv, reg)
api_surface_summary(audit$table)
eye_api_recommendation(audit)
```

The purpose is staged consolidation. Version 0.9 does not aggressively remove established interfaces.

<!-- BEGIN EYEPROCESS 0.9 API LIFECYCLE CLOSURE -->

## Packaged lifecycle closure

The 0.9 development surface is accompanied by a packaged lifecycle registry covering every currently exported API symbol. Classification is deliberately conservative: foundational front-door interfaces are `core`; orchestration and governance interfaces are `workflow`; specialist scientific interfaces are `advanced`; research-facing interfaces with explicit provisional contracts are `experimental`; and frontier or optional-engine interfaces may be `gated`.

Lifecycle status describes software-interface maturity and governance. It does not establish construct validity, empirical adequacy, device equivalence, or the scientific interpretation of gaze, pupil, response-time, sequence, or psychometric measures.

```{r lifecycle-closure}
lifecycle <- eyeprocess::eye_api_lifecycle()
inventory <- eyeprocess::eye_api_inventory()
audit <- eyeprocess::audit_eye_api(inventory, lifecycle)

eyeprocess::api_surface_summary(inventory)
audit
```

The current registry is intentionally frozen to the current exported surface. A future exported symbol that is absent from the registry continues to resolve as `unreviewed`; this makes lifecycle review an explicit release obligation rather than an inference from function naming.

<!-- END EYEPROCESS 0.9 API LIFECYCLE CLOSURE -->
