---
title: "Event-Detector Multiverse and Inference Robustness"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Event-Detector Multiverse and Inference Robustness}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 7,
  fig.height = 4.5
)
library(eyeprocess)
```

## Why event detection belongs in the analytical record

An event detector converts a continuous gaze stream into a discrete catalogue of
fixations, saccades, pursuits, and related events. That conversion is upstream of
AOI dwell, fixation count, time to first fixation (TTFF), revisits, transitions,
scanpaths, and fixation-locked pupil summaries. A substantive result can therefore
depend on the detector family or parameters even when the downstream model is
unchanged.

The detector-multiverse workflow asks a focused question:

> Would the substantive inference change under another defensible event-detection
> specification?

The workflow is:

```
raw gaze
  -> detector specification
  -> event catalogue
  -> AOI assignment
  -> derived features
  -> identical statistical model
  -> inference-stability report
```

This is not a search for a detector that produces a preferred p-value.

## When to use this workflow

Use detector sensitivity analysis when:

- fixation/saccade classification is upstream of the outcome used in the paper;
- more than one threshold or detector family is scientifically defensible;
- vendor events are being compared with research algorithms;
- a reviewer asks whether a gaze effect is detector-dependent;
- event fragmentation/merging could plausibly affect AOI metrics or scanpaths.

Do not construct arbitrary parameter grids and present them as universal
recommendations. Thresholds depend on sampling rate, precision, task, display,
head motion, pursuit content, preprocessing, and the scientific question.

## Define detector specifications

The core does not silently choose thresholds.

```{r}
ivt25 <- define_event_detector_spec(
  "ivt25",
  "ivt",
  velocity_threshold = 25,
  minimum_duration_ms = 60,
  maximum_gap_ms = 75,
  sampling_rate = 60,
  coordinate_unit = "degrees"
)

ivt35 <- define_event_detector_spec(
  "ivt35",
  "ivt",
  velocity_threshold = 35,
  minimum_duration_ms = 60,
  maximum_gap_ms = 75,
  sampling_rate = 60,
  coordinate_unit = "degrees"
)

idt_a <- define_event_detector_spec(
  "idt_A",
  "idt",
  dispersion_threshold = 1.2,
  minimum_duration_ms = 80,
  sampling_rate = 60,
  coordinate_unit = "degrees"
)

adaptive <- define_event_detector_spec(
  "adaptive",
  "adaptive_velocity",
  minimum_duration_ms = 60,
  maximum_gap_ms = 75,
  sampling_rate = 60,
  coordinate_unit = "degrees",
  parameters = list(
    noise_factor = 4,
    minimum_velocity_threshold = 20
  )
)

multiverse <- create_detector_multiverse(
  list(ivt25, ivt35, idt_a, adaptive)
)

multiverse$manifest[, c(
  "detector_id",
  "algorithm",
  "velocity_threshold",
  "dispersion_threshold",
  "minimum_duration_ms"
)]
```

The built-in adaptive detector is explicitly a **robust-MAD reference
implementation**. It is not labelled as REMoDNaV, Nyström-Holmqvist,
Engbert-Kliegl, or another named external algorithm.

## Synthetic 60 Hz worked example

The bundled generator is deliberately small, reproducible, and private-data free.
It creates a known positive disclosure-dwell effect.

```{r}
dat <- simulate_detector_multiverse_data(
  n_participants = 4,
  sampling_rate = 60,
  seed = 20260918
)

detected <- run_detector_multiverse(
  dat,
  multiverse,
  continue_on_error = TRUE
)

summarise_detector_events(detected)
```

The same raw gaze is processed independently for every specification.

## Compare event catalogues structurally

Different detectors do not necessarily produce the same number of rows, so
row-by-row comparison is inappropriate. `match_detected_events()` performs
deterministic one-to-one temporal matching using event overlap and onset
tolerance.

```{r}
agreement <- estimate_detector_agreement(detected)
agreement[, c(
  "detector_a",
  "detector_b",
  "matched_event_precision",
  "matched_event_recall",
  "f1",
  "mean_event_overlap"
)]
```

Without an independently justified reference catalogue, these quantities describe
**agreement**, not accuracy.

```{r detector-agreement, fig.cap="Pairwise temporal agreement among successful fixation catalogues."}
plot_detector_agreement(detected)
```

## Propagate detector choice into AOI features

AOI overlap defaults to `"error"`. If an event belongs to more than one AOI, the
analyst must choose a rule explicitly rather than receiving an undocumented
first-match assignment.

```{r}
features <- propagate_detector_to_aoi(
  detected,
  overlap = "error"
)
features <- propagate_detector_to_features(features)

head(features$features[, c(
  "detector_id",
  "participant_id",
  "trial_id",
  "condition_id",
  "aoi_id",
  "fixation_count",
  "dwell_time_ms",
  "ttff_ms",
  "revisits"
)])
```

Zero-event trials are retained. If a trial contains valid gaze but no target
fixation, count/dwell can legitimately be zero. If the trial contains no valid
gaze, those quantities remain missing rather than being converted to zero.

```{r feature-sensitivity, fig.cap="Synthetic disclosure-AOI dwell across detector specifications."}
plot_detector_feature_distributions(
  features,
  feature = "dwell_time_ms",
  aoi_id = "disclosure"
)
```

## Propagate into the same statistical model

The statistical engine is explicit. The example uses `stats::lm()` through
`engine = "stats_lm"`; mixed models can use `lme4_lmer`, and specialist
estimators can enter through the documented callback contract.

```{r}
model_spec <- list(
  engine = "stats_lm",
  formula = dwell_time_ms ~ condition_id + participant_id,
  outcome = "dwell_time_ms",
  aoi_id = "disclosure"
)

inference <- run_detector_inference_multiverse(
  features,
  model_spec,
  minimum_valid_fraction = 0.5
)

condition_term <- grep(
  "condition_id",
  inference$coefficients$term,
  value = TRUE
)[1]

stability <- assess_detector_inference_stability(
  inference,
  term = condition_term,
  substantive_threshold = 100,
  direction = "above"
)

stability
```

Formula models use `na.fail`. Missing predictors therefore cause an explicit
branch failure rather than an unnoticed reduction in N. Non-converged branches
are retained diagnostically and excluded from coefficient-stability summaries.

```{r coefficient-stability, fig.cap="Synthetic condition coefficient and confidence interval across detector specifications."}
plot_detector_coefficient_stability(
  inference,
  term = condition_term
)
```

## REMoDNaV is an external bridge, not a reimplementation

When REMoDNaV is the intended detector, request it explicitly:

```{r remodnav-spec, eval=FALSE}
remodnav <- define_event_detector_spec(
  "remodnav",
  "remodnav",
  minimum_duration_ms = 60,
  sampling_rate = 60,
  coordinate_unit = "degrees",
  parameters = list(
    noise_factor = 5
  )
)
```

`eyeprocess` calls the installed REMoDNaV command-line implementation. If the
executable is unavailable, that branch fails explicitly and no internal
approximation is substituted. Pixel data require an explicit `px2deg`
conversion factor; normalized coordinates must be converted before the bridge is
used.

## What to report

At minimum, report:

1. the primary detector and its rationale;
2. every evaluated detector family and parameter specification;
3. sampling rate and coordinate units;
4. thresholds, minimum duration, gap/merge rules, smoothing/filtering, and
   external implementation versions;
5. AOI definitions and the AOI-overlap rule;
6. quality rules applied before modelling;
7. successful, failed, and non-converged branches;
8. event-count/duration differences and temporal agreement;
9. sensitivity of the AOI features supporting the claim;
10. coefficient distributions, intervals, sign stability, and any explicitly
    declared substantive threshold;
11. source, preprocessing, detector, AOI, quality, model, and software
    provenance.

Do **not** summarize robustness by counting significant p-values.

## Generate a Markdown report

```{r}
report <- report_detector_multiverse(
  features,
  inference = inference,
  term = condition_term,
  substantive_threshold = 100
)

cat(substr(report, 1, 800), "...")
```

## Interpretation boundaries

Detector agreement does not establish event validity. Detector disagreement does
not establish which detector is correct. A stable coefficient across a narrow
multiverse does not imply stability to omitted preprocessing, AOI, quality, or
model choices.

For broader workflow sensitivity, see
`vignette("multiverse-sensitivity-and-decision-stability")`. For the canonical
event functions, see `?event_detector_multiverse`.

Methodological context includes Dar, Wagner, and Hanke (2021) on REMoDNaV and
recent event-detection evaluation work emphasizing signal/task dependence and
event-level matching rather than naive row pairing.
