---
title: "Frozen validation evidence programme"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Frozen validation evidence programme}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(eyeprocess)
```

Milestone #2 separates **software-validation evidence** from substantive construct validity. A reproducible plan fixes scenario families, sample sizes, missingness, perturbations, replications, and seeds before results are interpreted.

```{r}
plan <- eyeprocess_validation_plan(
  sample_size = c(250L, 750L), n_items = c(12L, 24L),
  missing_rate = c(0, .15), specification = c("correct", "misspecified"),
  replications = 20L, seed = 20260811L
)
head(expand_eyeprocess_validation_plan(plan))
```

Acceptance criteria are declared as reporting contracts rather than universal scientific thresholds.

```{r}
rules <- list(
  rmse = validation_acceptance_rule("rmse", "max", .20),
  failure = validation_acceptance_rule("failure_rate", "max", .05)
)
```

Frozen evidence objects contain integrity hashes and source-commit metadata. A passing release gate means only that the declared software-validation criteria were satisfied.
