---
title: "IRT information, scoring, and diagnostics"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{IRT information, scoring, and diagnostics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(eyeprocess)
```

The native IRT layer provides transparent mathematical utilities for common dichotomous and polytomous families. Estimation is not silently approximated when a specialized external engine is required.

```{r}
items <- data.frame(
  item_id = paste0("I", 1:8), a = seq(.8, 1.5, length.out = 8),
  b = seq(-1.5, 1.5, length.out = 8), c = 0, d = 1
)
info <- eyeprocess_irt_test_information(seq(-3, 3, by = .25), items)
head(info)
eyeprocess_irt_measurement_precision_profile(info$theta, items)
```

The module also supplies EAP/MAP/MLE score utilities, conditional uncertainty, bank targeting, residual fit, Q3/local-dependence summaries, and Infit/Outfit-style diagnostics. Person-fit quantities are model diagnostics; they must not be converted into claims about cheating, disengagement, diagnosis, or mental state.

Stan's current User Guide emphasizes explicit identification for IRT models and sparse encodings when response matrices are incomplete. These principles motivate `eyeprocess_irt_identification_audit()` and `eyeprocess_irt_sparse_design_audit()`.

Primary source: <https://mc-stan.org/docs/stan-users-guide/item-response-models.html>.
