---
title: "IRT recovery, SBC, and misspecification evidence"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{IRT recovery, SBC, and misspecification evidence}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(eyeprocess)
```

Simulation-based calibration (SBC) and parameter recovery answer different software-validation questions. Recovery checks whether fitted estimates reproduce known generating quantities under declared scenarios; SBC checks calibration of posterior computation under the declared generative model.

```{r}
design <- eyeprocess_irt_recovery_design(sample_size = 250L, n_items = 12L,
                                          missing_rate = c(0,.15), testlet_sd = c(0,.35),
                                          replications = 5L, seed = 20260811L)
head(design)
eyeprocess_irt_misspecification_suite()
```

Exact recovery fitting currently requires `mirt`; absence produces a gated result. SBC rank summaries reuse eyeprocess's rank-diagnostic infrastructure.

The SBC workflow follows the logic of Talts et al., *Validating Bayesian Inference Algorithms with Simulation-Based Calibration*: <https://arxiv.org/abs/1804.06788>.
