---
title: "M3 device transport, falsification controls, and sensor value"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{M3 device transport, falsification controls, and sensor value}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(eyeprocess)
```

## More sensors can create more failure modes

Pupil measurements can vary with device, sampling rate, gaze geometry, missingness and preprocessing. A four-channel model therefore needs falsification and transport diagnostics in addition to a richer likelihood. M3 exposes those concerns without asserting that a device label itself explains measurement differences.

## Device-dependent stress

```{r}
sim <- simulate_multimodal_m3(
  n_person = 80,
  n_item = 10,
  pupil_missingness = "device",
  seed = 20260815
)

audit <- audit_multimodal_m3_identifiability(sim)
audit$device
```

```{r, fig.width=7, fig.height=4}
plot(sim, type = "device")
plot(audit, type = "device")
```

These outputs identify device-specific shifts or availability patterns in the simulation. They do not establish measurement invariance. Empirical device transport requires repeated-device or appropriately linked data, explicit equivalence/invariance analysis and validation of the underlying pupil units and preprocessing semantics.

## Falsification controls

```{r}
neg <- multimodal_m3_negative_controls(sim, seed = 20260816)
neg$provenance
```

M3 includes pupil permutations within item and within person, phase randomization, a luminance-only pseudo-pupil, and an irrelevant synthetic pupil channel. These deliberately break different aspects of person/item alignment. They are tests of whether the analysis pipeline is too willing to manufacture process value; they are not causal interventions and not participant-behavior or misconduct detectors.

```{r, fig.width=8, fig.height=5}
plot(neg, type = "pupil_alignment")
```

## Sensor value is conditional

After an ablation lattice has been fitted, `multimodal_m3_process_information()` reports the response-target pupil gain per usable pupil observation and per analyst-supplied relative sensor cost. This can support design discussions about whether a pupil channel is worth collecting under a specific model and target.

```{r, eval=FALSE}
ab <- multimodal_m3_ablation(sim, chains = 4, parallel_chains = 4)
info <- multimodal_m3_process_information(ab, pupil_cost = 1.5)
info$sensor_value
plot(info, type = "sensor_value")
```

The calculation is intentionally labeled a **sensor value-of-information screen**, not an economic cost-effectiveness analysis. It does not account automatically for equipment depreciation, staff time, participant burden, calibration failures, or the value of non-response outcomes.

## Channel conflict

The same information object places response ELPD beside convergence diagnostics. This makes a useful failure mode visible: a channel can look incrementally predictive while substantially worsening computational stability, or can sharpen latent estimates without improving held-out response prediction. M3 keeps those dimensions separate so that a single improvement cannot hide a meaningful trade-off.
