---
title: "M3 recovery, identifiability, and failure-case validation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{M3 recovery, identifiability, and failure-case validation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(eyeprocess)
```

## Validation must include failure cases

A four-channel model is not validated because one synthetic dataset samples successfully. M3 therefore separates structural support, parameter recovery, posterior predictive behavior, MCMC diagnostics, missingness stress and falsification controls.

The simulator provides five pupil signal regimes: `informative`, `weak`, `null`, `redundant`, and `confounded`. The last three are particularly important. A framework that always reports pupil as useful even when it is null, duplicative of gaze, or generated only by nuisance variables has failed its scientific purpose.

```{r}
scenarios <- c("informative", "weak", "null", "redundant", "confounded")
examples <- lapply(seq_along(scenarios), function(k) {
  simulate_multimodal_m3(
    n_person = 40,
    n_item = 8,
    pupil_signal = scenarios[k],
    seed = 20260815 + k
  )
})

vapply(examples, function(z) audit_multimodal_m3_identifiability(z)$supported, logical(1))
```

Structural support is deliberately different from incremental information. A null pupil channel can vary and be identifiable as a measurement dimension while still adding no response-target value.

## Recovery programme

```{r, eval=FALSE}
rec <- multimodal_m3_recovery(
  reps = 20,
  pupil_signal = c("informative", "weak", "null", "redundant", "confounded"),
  pupil_missingness = c("mcar", "quality", "device"),
  n_person = 150,
  n_item = 15,
  seed = 20260815,
  fit_args = list(
    chains = 4,
    parallel_chains = 4,
    iter_warmup = 750,
    iter_sampling = 750,
    refresh = 0,
    init = 0
  )
)

rec$summary
plot(rec, type = "rmse")
plot(rec, type = "coverage")
plot(rec, type = "pupil")
```

The recovery table reports bias, RMSE, posterior SD, 95% interval coverage, maximum R-hat and minimum bulk ESS by parameter family and stress condition. A tiny replication count is only a software smoke test; publication evidence should use a predeclared grid large enough to assess Monte Carlo uncertainty.

## Identifiability boundaries

M3 fixes person latent means at zero, response discrimination at one, RT person loading at -1, and gaze/pupil person loadings at +1. Correlated person and item structures are estimated using Cholesky factors. The audit checks whether the observed person-item graph is connected for each required channel and whether each channel varies.

These constraints make the implemented parameterization estimable under supported designs; they do not constitute a proof of global identifiability for every extension or missing-data mechanism.

## What counts as a successful negative result?

For null or confound-only pupil simulations, a good framework should recover near-zero process signal within uncertainty, avoid systematic response-target gains under pupil ablation, and expose nuisance-driven apparent pupil effects through the negative-control and confound diagnostics. M3 therefore treats “pupil adds no clear defensible information” as a legitimate validation outcome.
