---
title: "Multidimensional, testlet, and latent-regression IRT"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multidimensional, testlet, and latent-regression IRT}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(eyeprocess)
```

Milestone #2 adds design and diagnostic support for multidimensional IRT without duplicating mature estimation engines.

```{r}
L <- matrix(c(1,0, 1,0, 0,1, 0,1, 1,0, 0,1), ncol = 2, byrow = TRUE)
loading_spec <- eyeprocess_mirt_loading_spec(paste0("I",1:6), L, c("accuracy","process"), simple_structure = TRUE)
eyeprocess_mirt_loading_audit(loading_spec, min_items_per_dimension = 2)
eyeprocess_mirt_information_matrix(c(0,0), c(1,.5))
```

Testlet declarations, directional information, latent-regression design matrices, and identification audits are native. Exact multidimensional, bifactor/two-tier, multiple-group, mixed, and polytomous estimation is delegated to `mirt` or `TAM` where requested.

Primary package sources: <https://cran.r-project.org/package=mirt> and <https://cran.r-project.org/package=TAM>.
