---
title: "Multiple-Response Items, Revisiting, and Local Dependence"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Multiple-Response Items, Revisiting, and Local Dependence}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(eyeprocess)
```

# Why preserve the response process?

Multiple-response items can contain substantially more information than a
single total or partial-credit score. A participant who selects `A + C` and a
participant who selects `B + D` may receive the same conventional score while
showing very different option-level response and visual-inspection patterns.
The 0.7 development layer therefore preserves response combinations before any
scoring rule is imposed.

```{r}
long <- data.frame(
  participant_id = rep(c("p1", "p2"), each = 4),
  item_id = "item1",
  option_id = rep(c("A", "B", "C", "D"), 2),
  selected = c(TRUE, FALSE, TRUE, FALSE,
               FALSE, TRUE, FALSE, TRUE)
)
encode_response_combinations(long)
```

# Option-level process evidence

When option AOIs are available, the analysis can retain fixation/dwell evidence
at exactly the same option level as selection. `fit_multiple_response_process_irt()`
provides a transparent crossed-logistic reference model and an explicit external
engine gate.

```{r, eval=FALSE}
fit <- fit_multiple_response_process_irt(
  option_trials,
  selected = "selected",
  theta = "theta",
  item = "item_id",
  option = "option_id",
  gaze = "option_dwell_ms",
  engine = "reference"
)
```

The reference model is **not** the MRM/MRM-LD likelihood of Zhou and Guo. It is
provided to establish the data contract, generate empirical diagnostics, and
support validation before an exact implementation is connected. For a validated
exact implementation, use `engine = "external"` and retain engine/version
provenance.

# Local dependence must be checked

Inter-option local dependence can invalidate an analysis that treats option
responses as conditionally independent. If residuals from the response model
are available, `audit_process_local_dependence()` provides a Q3-style pairwise
diagnostic. An aligned process-residual matrix can be supplied to ask whether
response and gaze residual dependence show the same pair structure.

```{r}
set.seed(1)
r <- matrix(rnorm(400), ncol = 4,
            dimnames = list(NULL, paste0("option", 1:4)))
p <- r + matrix(rnorm(400, sd = .3), ncol = 4)
ld <- audit_process_local_dependence(r, p)
head(ld$pairs)
plot(ld)
```

The threshold is descriptive. It is not a universal significance cutoff and
must be interpreted with the fitted model, item design, multiplicity, and a
simulation-calibrated null distribution.

# Revisiting as collateral evidence in cognitive diagnosis

Current process-data work also shows that response time and item revisiting can
be modeled alongside cognitive-diagnosis responses. The eyeprocess adapter keeps
mastery semantics anchored to the supplied Q-matrix and uses revisiting, RT, and
optional gaze variables as collateral process evidence.

```{r, eval=FALSE}
cdm <- fit_revisit_process_cdm(
  response_matrix = Y,
  q_matrix = Q,
  process_data = process_log,
  person_id = "participant_id",
  revisited = "revisit_count",
  rt = "response_time_ms",
  gaze = c("stem_dwell_ms", "option_transition_count")
)
```

A process association must not be interpreted as a diagnosis of motivation,
misconduct, or cognitive state. The appropriate scientific question is whether
the process channel improves validated measurement or classification under
pre-specified external/grouped validation.

# Validation requirements

Before either model family is promoted, include at least response/attribute
recovery, local-dependence misspecification, option sparsity, process-channel
ablation, negative controls, and held-person/item/session/device validation.
Use `irt_validation_spec()`, `stress_test_local_dependence()`,
`process_channel_ablation()`, and `grade_model_evidence()` to retain a common
evidence record.
