---
title: "Quality, Provenance, and Responsible Interpretation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quality, Provenance, and Responsible Interpretation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", eval = FALSE)
```

Eye and physiological signals are observations. They do not identify mental
states without a defensible measurement model and external validation.

```{r}
validate_eye_dataset(x)
audit_sampling_rate(x)
audit_signal_quality(x)
audit_pupil_quality(x)
audit_trial_coverage(x)
audit_missingness(x)
analysis_readiness(x)
interpretive_warnings()
```

The package does not automatically equate:

- fixation with attention;
- long dwell time with item difficulty;
- pupil dilation with cognitive load;
- rapid response with guessing;
- EDA with a uniquely identified emotional state;
- a statistical process component with effort or engagement.

## Reproducible records

```{r}
provenance_manifest(x)
write_provenance(x, "provenance.json", format = "json")
report_eye_dataset(x, "analysis-report.md", include_plots = TRUE)
```

Exploratory and confirmatory analyses should use separately declared feature,
preprocessing, AOI, exclusion, and model specifications.
