---
title: "Trial-Level Multilevel Gaze Mediation Preparation"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Trial-Level Multilevel Gaze Mediation Preparation}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

## Why keep trials?

Participant aggregation mixes trial-to-trial deviations with stable between-person differences. The mediation preparation API preserves the trial as the observational unit and creates explicit `X_within`, `X_between`, `M_within`, and `M_between` columns.

## Preparation

```{r, eval=FALSE}
prepared <- prepare_multilevel_mediation_data(
  trials,
  x_col = "ai_correct",
  mediator_col = "source_dwell",
  outcome_col = "correct_override",
  quality_col = "valid_fraction",
  minimum_quality = 0.80,
  quality_action = "flag"
)
```

Missing gaze remains missing. A genuine observed zero is retained and separately identified. Poor-quality trials are flagged rather than silently discarded.

## Serial mediators or moderators

```{r, eval=FALSE}
prepared <- add_multilevel_mediation_component(
  prepared,
  value_col = "trust",
  semantic = "mediator2",
  within_col = "M2_within",
  between_col = "M2_between"
)
```

The statistical backend should consume these prepared components rather than reimplement the decomposition.


## Cross-language provenance

R records excluded/input row positions using R's one-based convention; `eyeprocesspy` records Python's zero-based convention. The convention is stored in provenance so cross-language comparisons do not silently shift row identities.
